---
title: "Automated Valuation Model"
editor: visual
---

<<<<<<< HEAD
```{r} 
library(sf)
library(tidyverse)
library(tidycensus)
library(tigris)
library(scales)
library(patchwork)
library(RColorBrewer)
library(here)

#LOAD SPATIAL DATA

COMMERICAL_COR <- st_read("data/CommercialCord.shp")
redlining <- st_read("data/Redlining.shp")
NBHD <- st_read("data/NBHDS.shp")
PHL_LIMITS <-  st_read("data/City_Limits.shp")
=======
```{r}
#installing all possible packages
install.packages(c(
  "tidyverse",
  "tidycensus",
  "scales",
  "RColorBrewer",
  "sf",
  "tigris",
  "patchwork",
  "knitr",
  "ggplot2"
))

#loading all packages
library(tidyverse)
library(tidycensus)
library(scales)
library(RColorBrewer)
library(sf)
library(tidyverse)
library(tigris)
library(tidycensus)
library(scales)
library(patchwork)
library(knitr)
library(ggplot2)

#loading data: median hh income, those with bachelors degree attainment, and white population presence
pa_county_data <- get_acs(
  geography = "block group",
  state = "PA",
  county = "Philadelphia",
  variables = c( 
    MEDHHINC = "B19013_001",
    BACHDEG = "B15003_022",
    WHITE = "B03002_003",
    Tpop = "B01003_001",
    Pov = "B17021_002",
    Vacant = "B25004_001",
    Total_Housing = "B25001_001"
  ),
  year = 2024,
  survey = "acs5",
  output = "wide"
)



#calculating margin of error for each 
pa_county_data <- pa_county_data %>%
  mutate(
    MEDHHINC_MOE = (MEDHHINCM / MEDHHINCE) * 100,
    BACHDEG_MOE  = (BACHDEGM / BACHDEGE) * 100,
    WHITE_MOE    = (WHITEM / WHITEE) * 100,
    Tpop_MOE     = (TpopM / TpopE) * 100,
    Pov_MOE = (PovM/PovE)*100,
    Vacant_MOE = (VacantM/VacantE)*100,
    Total_HousingMOE = (Total_HousingM/Total_HousingE)*100
  )
>>>>>>> b2c34a70b032cbc88f98dbce8a89adbdc5700002




<<<<<<< HEAD
```



```{r} 

#COLUMNS
colnames(NBHD)
colnames(redlining)
colnames(COMMERICAL_COR)


NBHD <- NBHD %>%
  select(LISTNAME, geometry) %>%
  rename(Name = LISTNAME) 

# Redlining
redlining <- redlining %>%
  filter(METRO_NAME == "Philadelphia-Camden-Wilmington") %>%
  select(GEOID20, INTPTLAT20, INTPTLON20, EQINTER20, METRO_NAME, geometry) %>%
  rename(
    Lon   = INTPTLON20,
    Lat   = INTPTLAT20,
    Metro = METRO_NAME,
    Grade = EQINTER20  
  )

# COMMERICAL_COR 
COMMERICAL_COR <- COMMERICAL_COR %>%
  select(P_DIST, PT_ADD, ZIP, geometry) %>%
  rename(
    District = P_DIST,
    Address  = PT_ADD,
    Zip      = ZIP
  )



```



```{r} 

#Clipping each shp file to PHL  boundaries- ONLY FOR REDLINING
 # Transform to the SAME CRS (4326 = standard lat/lon)

PHL_LIMITS <- st_transform(PHL_LIMITS, crs = 4326)
redlining   <- st_transform(redlining, crs = 4326)
NBHD        <- st_transform(NBHD, crs = 4326)
COMMERICAL_COR <- st_transform(COMMERICAL_COR, crs = 4326)

# redlining to Philly boundary
redlining_phl <- st_intersection(redlining_phl, PHL_LIMITS)


#it worked?
nrow(redlining)      # 13,488
nrow(redlining_phl)  # should be smaller





=======
#categorizing for future filteration 
filter_pa_category <- pa_county_data %>%
  mutate(
    reliability = case_when(
      is.na(MEDHHINC_MOE) | is.infinite(MEDHHINC_MOE) |
      is.na(BACHDEG_MOE)  | is.infinite(BACHDEG_MOE)  |
      is.na(WHITE_MOE)    | is.infinite(WHITE_MOE)     |
      is.na(Tpop_MOE)     | is.infinite(Tpop_MOE)      |
      is.na(Pov_MOE)      | is.infinite(Pov_MOE)       |
      is.na(Vacant_MOE)   | is.infinite(Vacant_MOE)    |
      is.na(Total_HousingMOE) | is.infinite(Total_HousingMOE) ~ "Unreliable",
      MEDHHINC_MOE > 40 | BACHDEG_MOE > 40 | WHITE_MOE > 40 |
      Tpop_MOE > 40 | Pov_MOE > 40 | Vacant_MOE > 40 | 
      Total_HousingMOE > 40 ~ "Use with Caution",
      TRUE ~ "Reliable"
    )
  ) %>%
  drop_na()



head(good_data)

final_dataset <- filter_pa_category %>%
  mutate(
    BACH_PROP = (BACHDEGE/ TpopE) * 100,
    WHITE_PROP = (WHITEE/TpopE) *100,
    Pov_Prop = (PovE/TpopE) *100,
    Vacancy_Rate = (VacantE/Total_HousingE)*100
  )%>%
  select(GEOID,
         NAME,
         BACH_PROP, 
         WHITE_PROP,
         Pov_Prop,
         Vacancy_Rate,
         MEDHHINCE
  )

final_dataset
>>>>>>> b2c34a70b032cbc88f98dbce8a89adbdc5700002





```

<<<<<<< HEAD


```{r} 



# Export cleaned shp


# Rename to short names before writing
redlining_phl <- redlining_phl %>%
  rename(
    lat  = Lat,
    lon  = Lon,
    metro = Metro,
    grade = Grade
  )

st_write(redlining_phl, "data/redlining_phl.shp", append=FALSE)
st_write(NBHD,           "data/NBHD_clean.shp",       append=FALSE)
st_write(COMMERICAL_COR, "data/CommCor_clean.shp",    append=FALSE)
st_write(PHL_LIMITS,     "data/PHL_LIMITS_clean.shp", append=FALSE)







```



```{r} 









```



```{r} 









```


```{r} 









```


```{r} 









```
=======
\

\

\
\

\
\

\
\
>>>>>>> b2c34a70b032cbc88f98dbce8a89adbdc5700002
